---
title: Data report
output:
  html_document:
    number_sections: TRUE
    toc: TRUE
    code_folding: hide
params:
  df:  mtcars
  max_nb_distinct_hist:  10
  min_cor_to_draw:  0.4
---

<style>

body {
text-align: justify
font-size: 12px}

h1.title {
  font-size: 38px;
  color: DarkBlue;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}

</style>


## Generate general document

```{r, message = FALSE}


## Loading libraries
library(dplyr)
library(stringr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(corrplot)
library(devtools)
source_url("https://raw.github.com/lz01/NAwhere/master/NAwhere.R")
```

```{r}
## General functions
nb_na = function(x) {return(sum(is.na(x)))}
nb_nan = function(x) {return(sum(is.nan(x)))}
nb_inf = function(x) {return(sum(is.infinite(x)))}
nb_0 = function(x) {return(sum(x==0,na.rm=T))}
nb_1 = function(x) {return(sum(x==1,na.rm=T))}
## is_id checks if a word contains "id" not preceded and not followed by letters, which would indicate that it is an identifier column
is_id = function(x) {str_detect(str_to_lower(x),"(?<![a-z])id(?![a-z])")}
```

```{r}
df = params$df
```


# General description

Check that the input is a data.frame, or that it can be converted to a data.frame (if tibble or other)
```{r}
if(!is.data.frame(df))
{
	df = as.data.frame(df)
}
```

Input can be a data structure and a name for the response variable.
By default, no response variable, and all variables are treated the same.
If there is a variable response, add a part to the report for relation between response and each other variable



## Number of variables and rows
```{r}
p = ncol(df)
n = nrow(df)
```
There are `r p` variables in the data.
There are `r n` rows in the data.

## Are there NA values?

```{r, out.width = '\\textwidth'}
df %>% summarize_all(nb_na) %>% 
kable(caption = " Number of NA values") %>% 
kable_styling()
if(nb_na(df)>0) NAwhere(df)
```

## What are variable types?

```{r}
df %>% 
summarize_all(typeof) %>% 
kable(caption = "Variable type") %>% 
kable_styling()

```

## Are there factor variables?
```{r}
nb_factors = sum(apply(df,2,is.factor))
```

There are `r nb_factors` factor variables.
```{r, results = 'asis'}
if(nb_factors>0)
{
df %>% 
summarize_if(is.factor,is.factor) %>% 
kable(caption = "Factor") %>% 
kable_styling()
}
```


# Variables

## Individual variables
```{r, warning = FALSE, message = FALSE, results = "asis"}

for(i in 1:p)
{
	colname = names(df)[i]
	x = df[,i]
	
	## Programmatically add subheader
	cat("\n###",colname,"\n")
	
	nb_distinct = length(unique(x))
	y = c(typeof(x), nb_distinct, nb_na(x))
	vals = c("Type","Nb of distinct values","Nb of NAs")
	
	## hist_plot is a boolean that indicates whether we plot a histogram of the variable or not
	## We plot a red histogram for numeric variables
	## We plot an orange histogram for character variables with less than max_nb_distinct_hist distinct values (not to overplot)
	## We plot an aquamarine histogram for boolean variables
	
	
	hist_plot = FALSE
	
	## Case: numeric
	if(is.numeric(x)) 
	{
		y = c(y,nb_nan(x), nb_inf(x),nb_0(x),mean(x,na.rm=T),median(x,na.rm=T),min(x,na.rm=T),max(x,na.rm=T))
		vals = c(vals,"Nb of NaN","Nb of Inf","Nb of 0s","Mean","Median","Minimum","Maximum")
		gg = ggplot() + geom_histogram(aes(x), fill = "darkblue", col = "white") + xlab(colname)
		hist_plot = TRUE
	}
	
	## Case boolean:
	if(is.logical(x))
	{
		y = c(y,nb_1(x),nb_0(x))
		vals = c(vals,"Nb of TRUE","Nb of FALSE")
		gg = ggplot() + geom_histogram(aes(x), fill = "aquamarine", col = "white", stat = "count") + xlab(colname)
		hist_plot = TRUE
	}
	
	## Case character:
	## If it has less than max_nb_distinct_hist, we plot a histogram
	if(is.character(x) & nb_distinct <= params$max_nb_distinct_hist)
	{
		gg = ggplot() + geom_histogram(aes(x), fill = "orange3", col = "white", stat = "count") + xlab(colname)
		hist_plot = TRUE
	}
	
	
	
	#xtable::xtable(data.frame(vals,y), caption = colname)
	
	matrix(c(vals,y),ncol=2) %>% 
	kable() %>%
	kable_styling() %>%
	print() 
	
	if(hist_plot) 
	{
		print(gg)
	}
 	cat("\n")
}
```

## Relations between variables

### Correlation
Correlations between variables as a general plot

```{r, fig.width = 7}
#, out.width = '\\textwidth'
## We select numeric variables, that don't include the word id (we usually are not interested in id variables for correlations)
df_num = df %>% select_if(is.numeric) %>% select(current_vars()[!is_id(current_vars())])
cor_mat =  df_num %>% cor(use = "pairwise.complete.obs")

cor_mat %>% corrplot(method = "number", order = "hclust", col = colorRampPalette(c("blue","white","red"))(200))
#cor_mat %>% corrplot.mixed(lower = "number", upper = "shade")
```



### Scatter plots between correlated variables


```{r}
## We find the pairs of variables that have a correlation > min_cor_to_draw
high_cor = abs(cor_mat) > params$min_cor_to_draw
high_cor[upper.tri(high_cor, diag = TRUE)] = FALSE
pairs_ind = which(high_cor, arr.ind = TRUE)


## Draw scatterplots
for(i in 1:nrow(pairs_ind))
{
	## Indexes of the 2 variables to draw
	pair_id = pairs_ind[i,]
	## We use aes_string as we pass a string to aes
	gg_scatter = ggplot(df_num[,pair_id], aes_string(x = names(df_num)[pair_id[1]], y = names(df_num)[pair_id[2]])) + 
	geom_point(alpha = 0.5 , col = "darkblue") +
	geom_smooth(method = "lm")
	
	print(gg_scatter)
	cat("  \n  \n  \n")
}
```


### Boxplots between correlated variables



